The effects of prenatal corticosterone and thermal environment on numerical discrimination abilities in a lizard

### 2024-08-27 ms.qmd
Updates

---
docx: 
bibliography: "./bib/refs.bib"
csl: "./bib/behavioral-ecology-and-sociobiology.csl"
reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

## Introduction

The ability to discern between more and less quantities is a basic cognitive skill widely observed across the animal kingdom (**Stancher et al. 2015; Vonk and Beran 2012; Beran et al. 2014**), and is considered to be adaptive in different ecological contexts (**Nieder 2017, 2018**). Female lions (_Panthera leo_), for example, can use numerical information to assess the risk of confronting rival groups (**McComb et al. 1994**); but other animals can employ quantitative abilities for navigation (**Chitka and Geiger 1995**), mate selection (**Carazo et al. 2009**) or foraging (**Uller et al. 2003**). Given the pervasive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas. Some studies focus on the type of cues involved in quantitative competences (**REFs**), others on the context where animals may use these abilities (**REFs above**), and others in the neural mechanisms underlying quantitative and numerical discrimination (**REFs**). The later have demonstrated that non-verbal numerical representations activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant (**Nieder, 2005**). This link between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.

Animal's brain is particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014]. For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic]. These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016], but little is known about how exposure to GCs during development can affect quantitative abilities. Similarly, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019]. For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019]. However, the effects of early thermal environment on numerical abilities in other taxa remain unknown. In this sense, not only is it well-established that early-life GCs and thermal environment independently affect cognitive abilities, but these factors are also predicted to interact. In vertebrates, abrupt changes in temperature can lead to increased GC levels which, in turn, may elevate GCs in their offspring [@racic_effects_2020; @Crino_2023]. Simultaneously, these offspring are directly influenced by the same thermal environment, resulting in significant interactions that can impact cognitive development. As such, and given the ecological importance of quantitative abilities in nature [**Nieder 2017, 2018**], it is crucial to study the combined effects of early-life GCs and thermal environment on quantity and numerical discrimination. Here, we investigate the effects of prenatal exposure to corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and thermal environment on the ability of lizards to discriminate between different number of food items using the Common garden skink (_Lampropholis guichenoti)_ as a model species.

Traditionally, the study of numerical abilities has been mainly focused on mammals, birds, and fishes (**Agrillo and Bisazza 2014; Nieder 2017**), but recent research has shown that reptiles are also competent in numbers. Red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food (**Soldati et al. 2017**), Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small amounts off food in a spontaneous choice test (**Gazzola et al. 2018**), and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task (**Lin et al. 2021, 2024**). In lizards, __Podarcis siculus_ did not discriminate between different amounts of food in a spontaneous choice test (**Miletto-Petrazzini et al. 2017**), but were successful in a trained quantity discrimination task (**Miletto-Petrazzini et al. 2018**), rock lizards (_Iberolacerta cyreni_) chose bigger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021], and Gidgee skinks (_Egernia stokesii_) used numbers to select the larger amount of food in a spontaneous choice test [@szabo_spontaneous_2021]. Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed. However, the question of how early environment can affect numerical abilities remains open as, to our knowledge, there is no study investigating how early life conditions impact numerical discrimination in this group. Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020]. Our understanding of the effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].

In our study, we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design. After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they selected between different numbers of crickets. We hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task. Additionally, we predicted that increased CORT levels would impair numerical competence. Moreover, we expected the negative impact of CORT on numerical performance to be more pronounced in lizards incubated at lower temperatures compared to those in higher temperatures. In line with Weber's law [**REFs**], we further anticipated that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g. 3 VS 4), whereas event the lowest ratios (1 VS 4) could be challenging for those incubated with CORT.

@@

The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items. Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food. We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4. We never used more than four items because higher numbers are considered to be encoded by different regions of the brain [**REFs**]. The food consisted in frozen small crickets (*A. domestica*) that were thawed and dusted with calcium and multivitamin. We decided to use frozen crickets to avoid movement to affect lizards' behaviour. Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey. Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B). The side where the food was placed was changed every day to avoid any side bias during the trials. However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals. We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day. The crickets used during the experiment were all the same size, but to control for other quantitative traits 
(**Uller et al. 2003; Krusche et al. 2010; Abramson et al. 2011; Garland et al. 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015**), we changed the orientation of the crickets as shown in @fig-Methods D. In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).

@@

## Discussion
Our results indicate that lizards were not able to discriminate between different number of crickets in any of the tests, and that their performance did not vary across treatments. We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either. Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed. Similar findings were obtained on the interest for the larger number of food items, that did not differ from zero in any of the tests or groups. These results imply that _L. guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.

In out experiment, we employed a spontaneous choice test to investigate _L. guichenoti_'s numerical abilities. This test is considered a reliable method to assess numeosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination (**REFs**). However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance (**Agrillo and Bisazza, 2014**). Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice, and, as such, we do not consider motivation to be a factor affecting our results. Spontaneous choice tests are also based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one (**Agrillo and Bisazza, 2014**). Although the Optimal Foraging Theory predicts animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved (**REFs**). Nonetheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey. Finally, spontaneous choice tests are constraint by the researchers' ability to control for other quantitative cues that may correlate with numerosity (**Agrillo and Bisazza, 2014; Uller et al. 2003; Krusche et al. 2010; Abramson et al. 2011; Garland et al. 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015**). Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues. If lizards were using other cues exclusively for making their decision, we would still expect to see some type of discrimination between choices unless the cues were conflicting [**Cooper et al_Journal of Herpetology**]. Indeed, it could be that _L. guichenoti_ uses different quantitative cues at the same time in their foraging decisions, and, when conflicted, the preference towards one or the other varies between individuals, which may produce the same results we observed here. Alternatively, it could be that _L. guichenoti_ does not use numerical cues in their foraging decisions, and that they rely on other type of information to make their choices. Half of the studies showing numerosity in reptiles employed trained quantity discrimination tasks [**Soldati et al. 2017; Lin et al. 2021, 2024; Miletto-Petrazzini et al. 2018; Tomonaga et al. 2023 Animal Cognition**]. In the remaining studies, the relevant stimuli involved either vegetables [**Gazzola et al. 2018**; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs. 10)[@recio_prey_2021]. In contrast, _Podarcis siculus_ failed to discriminate between different numbers of _Musca domestica_ larvae in a spontaneous choice test, but they showed a clear preference towards bigger larvae in a similar task [**Miletto-Petrazzini et al. 2017**]; and they were able to learn to discriminate between small close numbers in a training procedure [**Miletto-Petrazzini et al. 2018**]. This suggest that numerosity may not be relevant in foraging contexts for reptiles that eat live prey. In this regard, numerical abilities may have evolved under different situations. [**EXAMPLE FISHES THAT NOT FORAGE BUT YES SOCIAL**]. Further studies should explore numerosity in _L. guichenoti_ with a different methodology and accross different contexts. **For instance: things about higher numbers (our paper), use different type of relevant stimuli (females, conspecifics? Carazo et al and other), use training procedures.**

- Treatment did not affect the discrimination capacity or the decision making
  - Other studies showing the effect of temperature on discrimination
  - Other studies about the effects of hormones on discrimination
For instance, mammals’ hippocampi contain more GCs receptors (GRs) than other regions of the brain (Siegel et al., 2006), which may be the main cause of overall impairments inspatial learning in rats but not in other cognitive domains (Haq et al., 2021).
  - Studies showing the effect of temperature and CORT on lizards' cognition
  - The 'two systems'
The second non-verbal system is thought to be an object tracking system (also known as ‘subitizing’) that yields relatively discrete representations. This system keeps track of a small number of items (originally set to 6 (REF. 32), but later restricted to 3 or 4 items) by assign-ing ‘markers’ (also known as ‘fingers’,‘pointers’,‘symbols’, 'tokens’ or ‘files’) to individual elements (**Nieder, 2005 Nature Reviews**)
Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al. 2009; Hyde 2011). **Are they encoded in different regions? Brin-nuclei-dependent effects of CORT and temp**

- Conclusions
Our results indicate that _L. guichenoti_ does not use numerical to make foraging decisions. This outcome, however, does not imply lack of numerical abilities, since other types of experimental approaches can yield different results. We also observed that lizards' performance was not different between treatments, which suggests that the early environment did not affect their numerical abilities. However, other tests and paradigmas should be used to test this further since both CORT and temperature can have different effects on different brain nuclei.

@@

#### Checking the models plots  
```{r, checking_model, out.width="70%", fig.align="center"}
# Chunk for plotting the model
## Initial plot
plot(model)
#
## pp_checks
pp_check(model, resp = "loglatency")
pp_check(model, resp = "choice")
pp_check(model, resp = "comparedinterest")
```

>>>>

---
title: "The effects of prenatal corticosterone and thermal environment on numerical discrimination abilities in a lizard"
docx: 
bibliography: "./bib/refs.bib"
csl: "./bib/behavioral-ecology-and-sociobiology.csl"
reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Chris Friesen^3^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^

^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au

## Abstract

## Significant statement

## Introduction  
The ability to discern between more and fewer quantities is a basic cognitive skill widely observed across the animal kingdom [@stancher2015numerical; @vonk2012bears; @beran2016capuchin]. It is considered adaptive in various ecological contexts, contributing to behaviors such as social decisions, mate selection, or foraging [@lucon2017individual; @nieder2018evolution]. For instance, female lions (_Panthera leo_) use numerical information to assess the risk of confronting rival groups [@mccomb1994roaring], while other animals use these abilities for navigating their environment [@chittka1995can], locate potential mates [@carazo2009quantity], or to make optimal foraging decisions [@uller2003salamanders]. Given the adaptive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas.

Numerical competences are often evaluated by considering the types of cues animals use [@plotnik2019elephants; @cooper2024tell], the ecological contexts in which these abilities are deployed [@mccomb1994roaring; @chittka1995can; @uller2003salamanders; @carazo2009quantity], and the neural mechanisms that underlie them [@nieder2005counting]. The later have demonstrated that non-verbal numerical representations in mammals activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant [@nieder2005counting]. This connection between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.

Animal's brain is particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014]. For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic]. These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016; @haq_prenatal_2021], but little is known about how exposure to GCs during development can affect quantitative abilities. At the same time, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019]. For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019]. However, the effects of early thermal environment on numerical abilities in other taxa remains largely unexplored.

Given that abrupt temperature changes can increase GC levels in vertebrates, leading to elevated GCs in offspring [@racic_effects_2020; @Crino_2023], it is essential to study the combined effects of early-life GCs and thermal conditions on cognitive development. Understanding how these factors influence quantitative abilities could reveal important insights into the ecological relevance of numerosity. To fill this gap, we used the common garden skink (_Lampropholis guichenoti_) as a model to test the effects of prenatal corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - exposure and thermal environment on numerical discrimination abilities.

Traditionally, research on numerical abilities has been mainly focused on mammals, birds, and fishes [@nieder2005counting; @agrillo2014spontaneous], but recent studies demonstrate that reptiles are also competent in numbers. For example, red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food [@soldati2017long], Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small food quantities in a spontaneous choice test [@gazzola2018continuous], and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task [@lin2021superior; @lin2024trained]. In lizards, __Podarcis siculus_ failed to discriminate between different number of food items in a spontaneous choice test [@miletto2017quantitative], but succeeded in a trained quantity discrimination task [@miletto2018quantity], whereas rock lizards (_Iberolacerta cyreni_) exhibited a clear preference for larger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021]. Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed.

At the same time, reptiles' cognition is known to be significantly influenced by the environment in which they develop. Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020]. The effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].  However, the combined influence of prenatal CORT and thermal environment remains unknown, particularly on numerical abilities.

Our study aimed to address this gap by manipulating CORT concentration in _L. guichenoti_ eggs and incubating them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design. After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they were presented with different numbers of crickets. Based in other studies that explored the cognitive effect of prenatal GCs or thermal early environment in reptiles and other taxa, we hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_quantity_2019]. Additionally, we predicted that increased CORT levels would impair numerical competence [@szuran_water_1994; @zhu_prenatal_2004], with more pronounced effects at lower temperatures temperatures. Consistent with Weber's law [@agrillo2014spontaneous], we expected that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g. 3 VS 4), whereas event the lowest ratios (e.g. 1 VS 4) could be challenging for those exposed to CORT and/or incubated at lower temperatures.

@@

The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items. Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food. We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4. We never used more than four items because higher numbers are considered to be represented by a different system [@nieder2005counting; @cantlon2009beyond; @hyde2011two]. The food consisted in frozen small crickets (*A. domestica*) that were thawed and dusted with calcium and multivitamin. We decided to use frozen crickets to avoid movement to affect lizards' behaviour. Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey. Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B). The side where the food was placed was changed every day to avoid any side bias during the trials. However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals. We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day. The crickets used during the experiment were all the same size, but to control for other quantitative traits [@agrillo2014spontaneous], we changed the orientation of the crickets as shown in @fig-Methods D. In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).

@@

## Discussion
Our results indicate that lizards were not able to discriminate between different number of crickets in any of the tests, and that their performance did not vary across treatments. We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either. Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed. These findings were further reinforced by similar results regarding the interest for the larger number of food items, that did not differ from zero in any of the tests or groups. Altogether, these findings suggest that _L. guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.

In our experiment, we employed a spontaneous choice test to investigate _L. guichenoti_'s numerical abilities. This test is considered a reliable method to assess numerosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination [@nieder2005counting; @agrillo2014spontaneous]. However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance [@agrillo2014spontaneous; @bisazza2014experim]. Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice. Therefore we do not consider motivation to be a factor affecting our results. Moreover, spontaneous choice tests are based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one [@agrillo2014spontaneous]. Although spontaneous choice tests employing food assume animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved [@agrillo2014spontaneous]. Nevertheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey.

Another consideration is the limitation of spontaneous choice tests to control for other quantitative cues that may correlate with numerosity [@agrillo2014spontaneous]. Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues. If that were the case, however, we would still see some type of discrimination between choices unless the cues were conflicting [@cooper2024tell]. Indeed, it is possible that _L. guichenoti_ uses multiple quantitative cues simoultaneously during foraging decisions, and, when those cues conflict, the preference for one cue over another might vary between individuals, potentially producing the results observed here. Alternatively, _L. guichenoti_ may not rely on numerical cues for foraging decisions, instead utilizing other type of information.

Interestingly, half of the studies demonstrating numerosity in reptiles employed trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained]. In the remaining successful studies, the relevant stimuli involved either vegetables [@gazzola2018continuous; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs. 10)[@recio_prey_2021]. However, when _Podarcis siculus_ were presented different numbers of _Musca domestica_ larvae in a spontaneous choice test, the lizards failed to discriminate between quantities [@miletto2017quantitative] despite showing high quantitative skills when size of prey was the main cue [@miletto2017quantitative], or when tested using training procedures [@miletto2018quantity]. Hence, it is possible that further studies exploring numerosity in _L. guichenoti_ using different methodologies or contexts, such as larger quantities [@nieder2005counting; @recio_prey_2021], inclusion of chemical stimuli [@carazo2009quantity; @plotnik2019elephants; @recio_prey_2021], using relevant stimuli like conspecifics [@carazo2009quantity; @lucon2017individual], or use training procedures [@agrillo2014spontaneous; @bisazza2014experim; @soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained], may yield more conclusive results.

We did not find any effect of prenatal conditions on _L. guichenoti_'s ability to use numerical information for making foraging decisions. We anticipated that lower incubation temperatures and exposure to CORT will impair lizards' performance, especially in those tests with higher ratios following Weber's Law [@agrillo2014spontaneous; @nieder2018evolution]. However, lizards' behaviour was not impacted by our treatments, suggesting that early environmental conditions do not affect numerical competence in this species. Our predictions were based in previous studies showing the effects of prenatal temperature and GCs on cognitive abilities in reptiles and other groups [@szuran_water_1994; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @lui2017chronic]. However, these effects can be species-specific, varying across cognitive domains and test types [@szuran_water_1994; @farrell_developmental_2015-learn; @bebus_associative_2016;@dayananda_incubation_2017; @abayarathna_effects_2020; @haq_prenatal_2021].

In this context, it is important to consider that both GCs and thermal environment can exert differential effects on distinct regions of the brain. For example, incubation temperature impacts mitochondrial activity and cell density in specific brain nuclei [@coomber_independent_1997; @amiel_effects_2017]. Similarly, impairments in spatial learning - but not in other learning tasks - in rats [@haq_prenatal_2021] are likely associated with the greater density of GCs receptors mammals' hippocampi compared to other regions of the brain [@siegel_2006]. While it remains unknown which regions of the reptiles' brain encode numerical information, our results suggest that relevant regions may be unaffected by the prenatal conditions tested here. However, researchers have proposed that non-verbal number representation is processed by two types of systems: one for small sets (≤ 4) that tracks and encodes objects as individual entities, and another and another for larger numerosities that operates through estimation processes [@nieder2005counting; @cantlon2009beyond; @hyde2011two]. Our data indicates that at least those brain regions involved in discrete numerical representation might be robust to the effect of prenatal CORT and temperature in _L. guichenoti_. Studies using higher quantities of food items may help elucidate the effects of prenatal conditions on _L. guichenoti_'s numerical abilities. Alternatively, it is also plausible that treatment effects are masked by the overall poor performance across most of the tests; which could be a consequence of the species' relying in other type of information for foraging rather than numerical information. Thus, future research should investigate the effects of prenatal CORT and thermal environment on numerical competence by employing trained quantity discrimination tasks [@soldati2017long; @miletto2018quantity; @lin2021superior; @tomonaga2023slowly; @lin2024trained].

In conclusion, our study is, to the best of our knowledge, the first to explore the impact of prenatal conditions on numerical abilities in reptiles. Our findings suggest that _L. guichenoti_ does not rely on numerical information for foraging decisions and that their performance does not vary across different prenatal treatments. This implies that early environmental conditions, such as incubation temperature and CORT exposure, may not affect numerosity in this species. However, it is critical to note that the methodologies employed in this study may not have fully captured the nuances of numerical cognition in _L. guichenoti_. Therefore, we recommend that future studies employ a broader array of tests and paradigms, including training-based tasks, larger sets of stimuli, and other sensory modalities, to more comprehensively assess numerical competence in this and related species. Such work will be pivotal in unraveling the effects of early environmental factors on cognitive abilities in reptiles and advancing our understanding of numerosity across taxa.

@@

#### Checking the model plots  
```{r, checking_model, out.width="70%", fig.align="center"}
# Chunk for plotting the model
## Initial plot
plot(model)
#
## pp_checks
pp_check(model, resp = "loglatency")
pp_check(model, resp = "choice")
pp_check(model, resp = "comparedinterest")
```
